---
title: "patientChat Model Benchmark Results"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{patientChat Model Benchmark Results}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

This article is a skeleton to report benchmark results for `patientChat` across
all models returned by `availableModels()`.

Benchmark task definition:

- Generate exactly 10 synthetic OMOP CDM v5.4 patients.
- Include diabetes in `condition_occurrence`.
- Include semaglutide in `drug_exposure`.
- Record elapsed generation time per model.

## 1) Run the benchmark script (manual step)

Use the benchmark test file:

```{r run-benchmark, eval = FALSE}
Sys.setenv(
  OPENAI_API_KEY = "<your-key>",
  PATIENTGENERATOR_RUN_MODEL_BENCHMARK = "true",
  PATIENTGENERATOR_MODEL_BENCHMARK_DIR = "/path/to/benchmark-output"
)

testthat::test_file("tests/testthat/test-patientChat-model-benchmark.R")
```

Expected output artifact:

- `/path/to/benchmark-output/patientChat_model_benchmark_results.csv`

## 2) Load benchmark CSV

Set `results_csv` to your generated file path and re-knit.

```{r load-results}
default_results_csv <- "vignettes/data/patientChat_model_benchmark_results_gpt5_main_2026-02-28.csv"
results_csv <- Sys.getenv("PATIENTGENERATOR_BENCHMARK_RESULTS_CSV", unset = default_results_csv)

results <- if (file.exists(results_csv)) {
  utils::read.csv(results_csv, stringsAsFactors = FALSE)
} else {
  message("Benchmark CSV not found. Set PATIENTGENERATOR_BENCHMARK_RESULTS_CSV and re-knit.")
  data.frame(
    model = character(),
    status = character(),
    elapsed_seconds = numeric(),
    person_count = integer(),
    output_file = character(),
    error = character(),
    stringsAsFactors = FALSE
  )
}
```

## Published snapshot (2026-02-28)

The repository currently includes one published benchmark run for core GPT-5 models.

```{r snapshot}
if (nrow(results) > 0) {
  successful <- results[results$status %in% c("success", "success_non_10"), , drop = FALSE]
  failed <- results[results$status == "error", , drop = FALSE]

  if (nrow(successful) > 0) {
    best <- successful[which.min(successful$elapsed_seconds), , drop = FALSE]
    cat(
      "Fastest successful model:", best$model,
      sprintf("(%.2f seconds).", best$elapsed_seconds), "\n"
    )
  }

  cat("Successful runs:", nrow(successful), "\n")
  cat("Failed runs:", nrow(failed), "\n")
}
```

## 3) Raw results table

```{r raw-table}
if (nrow(results) > 0) {
  knitr::kable(results, caption = "Raw benchmark output")
}
```

## 4) Successful runs ranked by speed

```{r ranking}
if (nrow(results) > 0) {
  successful <- results[results$status %in% c("success", "success_non_10"), , drop = FALSE]
  successful <- successful[order(successful$elapsed_seconds), , drop = FALSE]

  knitr::kable(successful, caption = "Successful runs sorted by elapsed_seconds")
}
```

## 5) Timing summary

```{r summary}
if (nrow(results) > 0) {
  successful <- results[results$status %in% c("success", "success_non_10"), , drop = FALSE]
  if (nrow(successful) > 0) {
    timing_summary <- data.frame(
      metric = c("n_success", "mean_seconds", "median_seconds", "min_seconds", "max_seconds"),
      value = c(
        nrow(successful),
        mean(successful$elapsed_seconds),
        stats::median(successful$elapsed_seconds),
        min(successful$elapsed_seconds),
        max(successful$elapsed_seconds)
      )
    )
    knitr::kable(timing_summary, caption = "Timing summary for successful runs")
  }
}
```

## 6) Timing plot

```{r plot}
if (nrow(results) > 0) {
  successful <- results[results$status %in% c("success", "success_non_10"), , drop = FALSE]
  if (nrow(successful) > 0) {
    ordered_idx <- order(successful$elapsed_seconds)
    barplot(
      successful$elapsed_seconds[ordered_idx],
      names.arg = successful$model[ordered_idx],
      las = 2,
      col = "#2E86AB",
      border = NA,
      ylab = "Elapsed seconds",
      main = "patientChat benchmark by model (lower is faster)"
    )
  }
}
```

## 7) Interpretation notes (fill after run)

Use this section to summarize:

1. Fastest stable model(s) for this cohort-generation task.
2. Models that returned errors or unstable outputs.
3. Recommended default model policy for package users.
