---
title: "Generate and Review Synthetic OMOP Patients"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Generate and Review Synthetic OMOP Patients}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

This article shows an end-to-end workflow with `patientGenerator`:

1. generate synthetic patients in OMOP CDM format with `patientChat`,
2. inspect and curate those patients with `patientDesigner`,
3. use Hecate concept search from the designer while editing concept IDs.

```{r setup}
library(patientGenerator)
```

## 1) Generate a synthetic cohort with patientChat

`patientChat` is an R6 class that prompts an LLM and returns schema-constrained JSON.

```{r chat-example, eval = FALSE}
# OPENAI_API_KEY must be available in your environment
generator <- patientChat$new(model = "gpt-5.2", echo = "none")

generator$prompt(
  paste(
    "Generate exactly 5 synthetic patients in OMOP CDM v5.4.",
    "Include one observation period and at least one condition occurrence per patient."
  )
)

# Save to default test-set location managed by the package
generator$save(name = "five-patients-demo")
```

### Optional: assist generation with a concept list

If you provide a concept list (`concept_id`, `concept_name`, `domain_id`, etc.),
`patientChat` can use it via the internal retrieval tool during generation.

```{r chat-codelist, eval = FALSE}
codelist <- readRDS(
  system.file("concept_sets", "ovarian_cancer_codelist.rds", package = "patientGenerator")
)

generator <- patientChat$new(
  model = "gpt-5.2",
  codelist_data = codelist,
  echo = "none"
)

generator$retrieveCodelist(concept_label = "ovarian", domain = "Condition")
```

## 2) Review and edit generated cohorts in patientDesigner

`patientDesigner()` starts a D3/Shiny application where you can load, review, and edit JSON test sets.

```{r designer-open, eval = FALSE}
patientDesigner()
```

Main capabilities in the app:

- browse and load available JSON test sets,
- edit `person`, `observation_period`, `condition_occurrence`, `drug_exposure`, and `measurement`,
- inspect timeline and raw table outputs,
- save updated cohorts back to JSON.

You can also target a specific folder:

```{r designer-path, eval = FALSE}
patientDesigner(path = "/path/to/my/testsets")
```

## 3) Search concept codes with Hecate from the designer

The concept search module in `patientDesigner` uses `hecateSearch()` under the hood.
This helps you find OMOP concept IDs while filling concept fields.

### Configure Hecate with environment variables

```{r hecate-env, eval = FALSE}
Sys.setenv(
  HECATE_BASE_URL = "https://your-hecate-server/api",
  HECATE_API_KEY = "your-api-key",
  HECATE_TIMEOUT_MS = "15000"
)
```

### Or configure with package options

```{r hecate-options, eval = FALSE}
options(patientgenerator.hecate = list(
  base_url = "https://your-hecate-server/api",
  timeout_ms = 15000,
  api_key = "your-api-key"
))
```

### Direct Hecate search example

```{r hecate-search, eval = FALSE}
res <- hecateSearch(
  query = "hypertension",
  domainId = "Condition",
  standardConcept = "S",
  limit = 20
)

head(res)
```

## Practical recommendations

- Start with small cohorts (1 to 5 patients) to validate your prompts.
- Keep prompts explicit about OMOP tables and required fields.
- Use `patientDesigner` for quality control before sharing test sets.
- Use Hecate filters (`domainId`, `standardConcept`) to reduce ambiguous concept matches.
